Constant-round Blind Classical Verification of Quantum Sampling

### 2021-05-31 introduction.tex
Sync Overleaf updates with GitHub

\section{Introduction}
Can quantum computation, with potential computational advantages that are intractable for classical computers,
be efficiently verified by classical means?
This seeming paradox has been one of the central problems in quantum complexity theory and delegation of quantum computation~\cite{web:Aaronson}.
From a philosophical point of view, this question is also known to have a fascinating connection to the \emph{falsifiability} of quantum mechanics in the high complexity regime~\cite{survey:AV12}.

A complexity-theoretic formulation of this problem by Gottesman in 2004~\cite{web:Aaronson} asks about the possibility for an efficient classical verifier (a $\BPP$ machine) to verify the output of an
efficient quantum prover (a $\BQP$ machine).
In the absence of techniques for directly tackling this question, earlier feasibility results on this problem have been focusing on two weaker formulations.
The first type of feasibility results (e.g.,~\cite{BFK09,arXiv:ABOEM17,FK17,mf16}) considers the case where the $\BPP$ verifier is equipped with limited quantum power.
The second type of feasibility results (e.g,~\cite{Nat:RUV13, CGJV19, Gheorghiu_2015, HPF15})
considers a $\BPP$ verifier interacting with at least two entangled, non-communicating quantum provers.
In a recent breakthrough, Mahadev~\cite{FOCS:Mahadev18a} proposed the first protocol of Classical Verification of Quantum Computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
The technique invented therein has inspired many  subsequent developments of CVQC protocols with improved parameters and functionality (e.g.,~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}).
On the other side, there are known complexity-theoretic limitations on the feasibility of blind CVQC in the information-theoretical setting (e.g.~\cite{aaronson_et_al:LIPIcs:2019:10582}).

With the newly developed techniques, we revisit the classical verification of quantum computation problems from both a philosophical and a practical point of view.
We first observe that the \emph{sampling} version of $\BQP$ (e.g., the class $\SampBQP$ formulated by Aaronson~\cite{aaronson_2013}) might be a more appropriate notion to serve the purpose of the original problem.
Philosophically, the outcomes of quantum mechanical experiments are usually samples or statistical information, which is well demonstrated in the famous double-slit experiment.
Moreover, a lot of quantum algorithms from Shor's~\cite{Shor} and Grover's~\cite{Grover} algorithms to some recent developments in machine learning and optimization (e.g.~\cite{brando_et_al:LIPIcs:2019:10603, AGGW17,pmlr-v97-li19b}) contain a significant quantum sampling component.
The fact that almost all quantum supremacy tasks (e.g.,~\cite{Boson, IQP, nature-google}) are sampling ones also suggests the relevance of delegation for quantum sampling problems.

It is worthwhile noting that there is a simple reduction of the delegation of \emph{classical} sampling problems to decision ones: for example, the verifier can fix a seed for pseudo-randomness, send it to the prover, and then the sampling outcome can be computed  bit-by-bit under this pseudo-random seed in a deterministic way. As such, classical literature of delegation of computation primarily focuses on the delegation of decision problems. 
Unfortunately, it is unclear how to make this de-randomization trick work for the delegation of $\SampBQP$, 
as randomness in quantum computation inherently comes from quantum mechanics rather than classical probability theory. 
Thus, it seems that the delegation of $\SampBQP$ needs to take a different technical route.

The first and important step in the case of $\SampBQP$ is to identify an appropriate definition of corresponding CVQC protocols. In contrast to the single-bit $\mathrm{Accept/Reject}$ information for the decision $\BQP$, any $\SampBQP$ problem needs to always output a sample of the desired distribution.
In the context of CVQC protocols, it means that the protocol should allow the verifier to generate the desired output sample whenever the protocol accepts. Of course, when the prover cheats, the verifier will reject and no further output is required in that case.  
Intuitively, the soundness should require that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interacting with a malicious prover.

A natural attempt is to directly apply Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} for $\SampBQP$.  Note that Mahadev's protocol consists of the so-called testing round, where the prover's behavior is tested,  and the Hadamard round, where measurement outcomes are generated. 
Indeed, some nice features of the Hadamard round,  e.g., allowing X-Z measurements on any qubit,  suggest the feasibility of generating  multi-bit measurement outcomes with the protocol. 
However, the requirement that the verifier needs to output a sample whenever accepting also exposes one drawback of Mahadev's protocol: while it is possible, as we will show, to generate desired output samples in the Hadamard round, 
it is however unclear how to generate such a sample when the protocol accepts in the testing round, since the sampling information stored in the quantum state will be disturbed by the testing step. 
Note that this difficulty is essentially due to the different requirements between the definitions of CVQC protocols for the decision and the sampling problems.

Another desirable property of CVQC protocols is the \emph{blindness} where the prover cannot distinguish the particular computation in the protocol from another one of the same size, and hence is blind about the client's input.
Historically, blindness has been achieved in the weaker formulations of CVQC based on various techniques: e.g., the measurement-based quantum computation exploited in~\cite{BFK09}, the quantum authentication scheme exploited in~\cite{arXiv:ABOEM17}, and the self-testing technique exploited in~\cite{Nat:RUV13}.
Moreover, the blindness property is known to be helpful to establish the verifiability of CVQC protocols. However, this is never an easy task.
See for example the significant amount of effort to add verifiability to blind computation protocol in~\cite{FK17}.
It is also known that another important primitive called the Quantum Fully Homomorphic Encryption (QFHE)~\cite{BJ15, DSS16, LC18, NS18, OTF18, mahadev_qfhe} should be helpful for establishing blindness: this is intuitive since QFHE allows fully homomorphic operations on encrypted quantum data. 
Indeed, in another paper~\cite{mahadev_qfhe}, Mahadev constructed the first leveled QFHE based on similar techniques and computational assumptions from~\cite{FOCS:Mahadev18a}.
The constructed QFHE automatically implies a blind CVQC protocol, however, without verifiability.
Extending this protocol with verifiability seems challenging as hinted by failed attempts in Section 2.2.2 of Mahadev's Ph.D. thesis~\cite{mahadev_2018}.

In fact, most existing blind and/or verifiable protocols for delegation of quantum computation require a notable amount of effort in achieving each property respectively. 
The only successful CVQC protocol~\cite{FOCS:GheVid19} of achieving both so far applies Mahadev's technique~\cite{FOCS:Mahadev18a} to the measurement-based quantum computation, whereas the analysis is still very specific to the construction.
Could there be a \emph{generic} way to achieve blindness and verifiability for CVQC protocols at the same time, say by leveraging the state-of-the-art CVQC~\cite{FOCS:Mahadev18a} and QFHE~\cite{mahadev_qfhe} protocols?

>>>>

\section{Introduction}

Can quantum computation, with potential computational advantages that are intractable for classical computers,
be efficiently verified by classical means?
This problem has been a major open problems in quantum complexity theory and delegation of quantum computation~\cite{web:Aaronson}. A complexity-theoretic formulation of this problem by Gottesman in 2004~\cite{web:Aaronson} asks about the possibility for an efficient classical verifier (a $\BPP$ machine) to verify the output of an
efficient quantum prover (a $\BQP$ machine).
In the absence of techniques for directly tackling this question, earlier feasibility results on this problem have been focusing on two weaker formulations.
The first type of feasibility results (e.g.,~\cite{BFK09,arXiv:ABOEM17,FK17,mf16}) considers the case where the verifier is equipped with limited quantum power.
The second type of feasibility results (e.g,~\cite{Nat:RUV13, CGJV19, Gheorghiu_2015, HPF15})
considers a $\BPP$ verifier interacting with at least two entangled, non-communicating quantum provers.

Recently, the problem is resolved by a breakthrough result of Mahadev~\cite{FOCS:Mahadev18a}, who constructed the first Classical Verification of Quantum Computation (CVQC) protocol for $\BQP$, where an efficient classical ($\BPP$) verifier can interact with an efficient quantum ($\BQP$) prover to verify any $\BQP$ language. Soundness of Mahadev's protocol  is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
%In a recent breakthrough, Mahadev~\cite{FOCS:Mahadev18a} proposed the first protocol of Classical Verification of Quantum Computation (CVQC) whose soundness is based on a widely recognized computational assumption that the learning with error (LWE) problem~\cite{JACM:Regev09} is hard for $\BQP$ machines.
The technique invented therein has inspired many subsequent developments of CVQC protocols with improved parameters and functionality. For example, Mahadev's protocol has a large constant soundness error. The works of ~\cite{arXiv:AlaChiHun19,arXiv:ChiaChungYam19} use parallel repetition to achieve a negligible soundness error. As another example, the work of ~\cite{FOCS:GheVid19} extends Mahadev's techniques in an involved way to obtain a CVQC protocol with an additional blindness property.

%For example,~\cite{arXiv:AlaChiHun19,arXiv:ChiaChungYam19}  
%
%(e.g.,~\cite{FOCS:GheVid19,arXiv:AlaChiHun19,arXiv:ChiaChungYam19}).

In this work, we make two more contributions to this exciting line of research. First, we observe that the literature has mostly restricted the attention to delegation of \emph{decision} problems (i.e., $\BQP$). Motivated by the intrinsic randomness of quantum computation and the sampling nature of many quantum applications, we initiate the study of CVQC for quantum \emph{sampling} problems. Second, we further investigate the desirable \emph{blindness} property and construct the first \emph{constant-round} blind CVQC protocols. We elaborate on our contributions in \Cref{subsection:sampling} and \ref{subsection:blind}, respectively.

% On the other side, there are known complexity-theoretic limitations on the feasibility of blind CVQC in the information-theoretical setting (e.g.~\cite{aaronson_et_al:LIPIcs:2019:10582}).

% With the newly developed techniques, we revisit the classical verification of quantum computation problems from both a philosophical and a practical point of view.

\subsection{CVQC for Quantum Sampling Problems} \label{subsection:sampling}

We initiate the study of CVQC for quantum sampling problem, which we believe is highly desirable and natural for delegation of quantum computation. In general, a round of quantum computation can be described as applying a quantum circuit and measuring the output qubits in computational basis. Due to its intrinsic randomness, the output is randomized and described by a distribution. 
%To verifiably delegate general quantum computation, it is natural to require that the output is close to the correct distribution when the verifier accepts. 
Thus, if a classical verifier want to utilize the full power of a quantum machine, the ability to get a verifiable sample from the quantum circuit's output distribution is desirable.  On a more concrete level, quantum algorithms like Shor's algorithm~\cite{Shor} has a significant quantum sampling component, and the recent quantum supremacy proposals (e.g.,~\cite{Boson, IQP, nature-google}) are built around sampling tasks, suggesting the importance of sampling in quantum computation.

%The previous works following Madadev's construction has been focusing on decision problems. In this work, we proposed and constructed a extension of Mahadev's protocol to \emph{sampling} problems, where the classical verifier get a sample from a desired distribution from the quantum prover. More formally, we seek a classical verifiable delegation of the class $\SampBQP$ proposed by Aaronson~\cite{aaronson_2013}.  Such an extension to sampling problems is desirable on a fundamental level; since we can always measure all qubits at the end of a quantum circuit, every quantum circuit naturally gives some distribution to sample from. Thus, if a classical verifier want to utilize the full power of a quantum machine, the ability to get a verifiable sample from the quantum circuit's output distribution is desired.  On a more concrete level, quantum algorithms like Shor's algorithm~\cite{Shor} has a significant quantum sampling component, and the recent quantum supremacy proposals (e.g.,~\cite{Boson, IQP, nature-google}) are built around sampling tasks, suggesting the importance of sampling in quantum computation.

% We first observe that the \emph{sampling} version of $\BQP$ (e.g., the class $\SampBQP$ formulated by Aaronson~\cite{aaronson_2013}) might be a more appropriate notion to serve the purpose of the original problem.

% Moreover, a lot of quantum algorithms from Shor's~\cite{Shor} and Grover's~\cite{Grover} algorithms to some recent developments in machine learning and optimization (e.g.~\cite{brando_et_al:LIPIcs:2019:10603, AGGW17,pmlr-v97-li19b}) contain a significant quantum sampling component.
% The fact that almost all quantum supremacy tasks (e.g.,~\cite{Boson, IQP, nature-google}) are sampling ones also suggests the relevance of delegation for quantum sampling problems.

It is worth noting that the difficulty of extending the delegation of decision problem to the delegation of sampling problems is quantum-specific. This is because 
there is a simple reduction from the delegation of \emph{classical} sampling problems to decision ones:  the verifier can sample and fix the random seed of the computation, 
%\footnote{One can use pseudorandomness, e.g., generated by a PRG seed, if verifier's efficiency is required, as typically in the classical delegation of computation literature.}, 
which makes the computation deterministic. Then, the verifier can delegate the output of the computation bit-by-bit as decision problems. However, this derandomization trick does not work in the quantum setting due to its intrinsic randomness.

%fix a seed for pseudo-randomness, send it to the prover, and then the sampling outcome can be computed  bit-by-bit under this pseudo-random seed in a deterministic way. Therefore, classical literature of delegation of computation primarily focuses on the delegation of decision problems. 
%However, it is unclear how to make this de-randomization trick work for the delegation of $\SampBQP$,  as randomness in quantum computation comes from inherent nature of quantum mechanics rather than some random seeds. 
%Thus, it seems that the delegation of $\SampBQP$ needs to take a different technical route.

% The first and important step in the case of $\SampBQP$ is to identify an appropriate definition of corresponding CVQC protocols.

\vspace{-3pt}

\paragraph{Our Contribution.}
As the first step to formalize CVQC for quantum sampling problems, we formalize efficiently computable quantum sampling problems with the complexity class $\SampBQP$ introduced by Aaronson~\cite{aaronson_2013}.  
%
%Formally, 
$\SampBQP$ consists of sampling problems $(D_x)_{x\in\zo^*}$ that can be approximately sampled by a $\BQP$ machine (See Section~\ref{sec:samp_definition} for the formal definition). We consider CVQC for a $\SampBQP$ problem $(D_x)_{x\in\zo^*}$ where a classical $\BPP$ verifier delegate the computation of a sample $z\leftarrow D_x$ for some input $x$ to a quantum $\BQP$ prover. Completeness requires that when the prover is honest, the verifier should accept with high probability and learn a correct sample $z\leftarrow D_x$. For soundness, intuitively, the verifier should not accept and output a sample with incorrect distribution  when interacting with a malicious prover. We formalize the soundness by a strong \emph{simulation-based} definition, (\Cref{dfn:stats-secure-proto-sampbqp}),
where we require that the joint distribution $(d,z)$ of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in  either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from the desired distribution $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.\hannote{justify?}
%
%To the best of our knowledge, we are the first to formally define delegation of quantum sampling problems.

%and almost never accept and output from an incorrect distribution when interacting with a malicious prover. We formalize this by a strong \emph{simulation-based} definition (Section~\ref{sec:samp_definition}), where we require that the joint distribution of the decision bit $d \in \set{\Acc, \Rej}$ and the output $z$ (which is $\bot$ when $d = \Rej$) is $\eps$-close (in  either statistical or computational sense) to an ``ideal distribution'' $(d,z_{ideal})$, where $z_{ideal}$ is sampled from the desired distribution $D_x$ when $d = \Acc$ and set to $\bot$ when $d = \Rej$.

As our main result, we construct a constant-round CVQC protocol for $\SampBQP$, based on the quantum LWE (QLWE) assumption that the learning-with-error problem is hard for BQP machines. 
\begin{theorem}[informal] \label{thm:qpip0-informal}
Assuming the QLWE assumption, there exists a four-message CVQC protocol for all sampling problems in $\SampBQP$ with computational soundness and negligible completeness error.
\end{theorem}

% To the best of our knowledge, we are the first to formally define the delegation of quantum sampling problems.

% We note that while several constructions in the relaxed models (e.g., verifiable blind computation~\cite{FK17}) can be naturally generalized to delegate quantum sampling problems and allow the verifier to learn multi-bit outputs, it seems non-trivial to show that these constructions achieve the simulation-based soundness property we defined. Proving the soundness of these constructions for delegating $\SampBQP$ is an interesting open question.

% The requirement that the verifier needs to output a sample when accepts makes a direct application of Mahadev's protocol infeasible for $\SampBQP$. As a result, in order to construct a $\QPIP_0$ protocol for $\SampBQP$,  one needs to (1) construct a $\QPIP_1$ protocol for $\SampBQP$ with very small completeness and soundness errors, and (2) amend Mahadev's construction so that the protocol will generate the desired sample whenever it accepts. 
% Our contribution is a solution to address the above two technical challenges.

% In contrast to the single-bit $\mathrm{Accept/Reject}$ information for the decision $\BQP$, any $\SampBQP$ problem needs to always output a sample of the desired distribution.
% In the context of CVQC protocols, it means that the protocol should allow the verifier to generate the desired output sample whenever the protocol accepts. Of course, when the prover cheats, the verifier will reject and no further output is required in that case.  
% Intuitively, the soundness should require that the verifier $V$ should never be ``cheated'' to accept and output an incorrect sample even when interacting with a malicious prover.

% 
% Indeed, some nice features of the Hadamard round,  e.g., allowing X-Z measurements on any qubit,  suggest the feasibility of generating  multi-bit measurement outcomes with the protocol.

%\paragraph{Techniques.}

The construction of our CVQC protocol follows the blueprint of Mahadev's construction~\cite{FOCS:Mahadev18a}. However, there are several obstacles we need to overcome along the way. To explains the obstacles and our ideas, we first present a high-level overview of Mahadev's protocol.

%Our protocol is constructed by generalizing Mahadev's protocol to ha

\paragraph{Overview of Mahadev's Protocol.}
%
Following~\cite{FOCS:Mahadev18a}, we define $\QPIP_{\tau}$ as classes of interactive proof systems between an (almost) classical verifier and a quantum prover, where the classical verifier has limited quantum computational capability, formalized as possessing $\tau$-qubit quantum memory. %\footnote{Intuitively, the size of local quantum memory limits the size of qubits that a general (entangled) quantum operation can be operated on at the same time. However, the verifier is still able to operate on all the qubits in a streaming fashion: i.e., receive the qubits sent from the prover and discard some local qubits in the case of overflow.}. 
A formal definition is given in Section~\ref{sec:qpip_def}.

At a high-level, the heart of Mahadev's protocol is a measurement protocol $\PiMeasure$ that can compile an one-round $\QPIP_1$ protocol with special properties to a $\QPIP_0$ protocol. Note that in a $\QPIP_1$ protocol, the verifier with one-qubit memory can only measure the prover's quantum message qubit by qubit. Informally, the measurement protocol $\PiMeasure$ allows a $\BQP$ prover to ``commit to'' a quantum state $\rho$ and a classical verifier to choose an $X$ or $Z$ measurement to apply to each qubit of $\rho$ such that the verifier can learn the resulting measurement outcome.

Thus, if an (one-round) $\QPIP_1$ verifier only applies $X$ or $Z$ measurement to the prover's quantum message, we can use the measurement protocol $\PiMeasure$ to turn the $\QPIP_1$ protocol into a $\QPIP_0$ protocol in a natural way. However, one restriction here is that the verifier's measurement choices need to be determined at the beginning (i.e., cannot depend adaptively on the intermediate measurement outcome).

Furthermore, in $\PiMeasure$, the verifier chooses to run a ``\emph{testing}'' round or a ``\emph{Hadamard}'' round with $1/2$ probability, respectively. Informally, the testing round is used to ``test'' the commitment of $\rho$, and the Hadamard round is used to learn the measurement outcome. (See \Cref{proto:urmila4} for further details about the measurement protocol $\PiMeasure$.) Another limitation here is that in the testing round, the verifier may accept the test of the commitment without learning any measurement result.

In~\cite{FOCS:Mahadev18a}, Mahadev's CVQC protocol for $\BQP$ is constructed by applying her measurement protocol to the one-round $\QPIP_1$ protocol of~\cite{PhysRevA.93.022326, mf16}, which has the desired property that the verifier only performs non-adaptive $X/Z$ measurement to the prover's quantum message. The fact that the verifier does not learn the measurement outcome in the testing round is not an issue here since the verifier can simply accept when the test is passed (at the cost of a constant soundness error).

\paragraph{Overview of Our Construction.}
As mentioned, our construction of CVQC protocol for $\SampBQP$ follows the blueprint of Mahadev's construction. Namely, we proceed in the following two steps: 1. construct a $\QPIP_1$ protocol for $\SampBQP$ with required special property, and 2. compile the $\QPIP_1$ protocol using $\PiMeasure$ to get the desired $\QPIP_0$ protocol. Both steps requires non-trivial works, but it turns out that the second step is the main technical challenge. At a high-level, the reason is the above-mentioned issue that the verifier does not learn the measurement outcome in the testing round. While this is not a problem for the case of decision problems (where the verifier only needs to decide to accept or reject), for the case of sampling problems, the verifier needs to produce an output sample when accepts, but there seems no way to produce the output for the verifier without learning the measurement outcome.   
%
We discuss both steps in turn as follows.

%As we shall see, both steps requires non-trivial works, and the second step is the main technical challenge.

\emph{Construct a $\QPIP_1$ protocol for $\SampBQP$ with required special property}: Interestingly, while the notion of delegation for quantum sampling problem is not explicitly formalized in their work, Hayashi and Morimae~\cite{hayashi2015verifiable} constructed an one-round $\QPIP_1$ protocol that can delegate quantum sampling problem and achieve our notion of completeness and soundness\footnote{They did not prove our notion of soundness for their construction, but it is not hard to prove its soundness based on their analysis.}. Furthermore, their protocol has information-theoretic security and additionally achieve the blindness property. However, in their protocol, the computation is performed by the verifier using measurement-based quantum computation (MBQC)\footnote{In more detail, the prover of their protocol is required to send multiple copies of the graph states to the verifier (qubit by qubit). The verifier tests the received supposedly graph states using cut-and-choose and perform the computation using MBQC.}, and hence the verifier needs to perform adaptive measurement choices. Therefore, we cannot rely on their $\QPIP_1$ protocol for $\SampBQP$.

Instead, we construct the desired $\QPIP_1$ protocol for $\SampBQP$ by generalizing the approach of local Hamiltonian reduction used in~\cite{PhysRevA.93.022326, mf16} to verify $\SampBQP$. Doing so requires the combination of several existing techniques from different context with some new ideas. For example, to handle $\SampBQP$, we need to prove lower bound on the spectral gap of the reduced local Hamiltonian instance, which is reminiscent to the simulation of quantum circuits by adiabatic quantum computation~\cite{adiabatic}. To achieve soundness, we introduce a cut-and-choose procedure and analyze it using de Finetti theorem in a  way similar to~\cite{takeuchi2018verification,hayashi2015verifiable}. See Section~\ref{sec:qpip1} for detailed discussions.

\emph{$\diamond$ Compile the $\QPIP_1$ protocol using $\PiMeasure$}: We now discuss how to use Mahadev's measurement protocol to compile the above $\QPIP_1$ protocol for $\SampBQP$ to a $\QPIP_0$ protocol. As mentioned, a major issue we need to address in Mahadev's original construction is that when the verifier $V$ chooses to run a testing round, $V$ does not learn an output sample when it accepts.  %(which happens with probability $1/2$),

Specifically, let $\PiNaive$ be an ``intermediate'' $\QPIP_0$ protocol obtained by applying Mahadev's compilation to the above $\QPIP_1$ protocol. In such a protocol, when the verifier $V$ chooses to run the Hadamard round, it could learn a measurement outcome from the measurement protocol and be able to run the $\QPIP_1$ verifier to generate a decision and an output sample when accepts. However, when  $V$ chooses to run the testing round, it only decides to accept/reject without being able to output a sample.

 A natural idea to fix the issue is to execute multiple copies of $\PiNaive$ in parallel\footnote{It is also reasonable to consider sequential repetition, but we consider parallel repetition for its advantage of preserving the round complexity.}, and to choose a random copy to run the Hadamard round to generate an output sample and use all the remaining copies to run the testing round. The verifier accepts only when all executions accept and outputs the sample from the Hadamard round. We call this protocol $\PiSampZ$.

Clearly from the construction, the verifier now can output a sample when it decides to accept, and output a correct sample when interacting with an honest prover (completeness). The challenge is to show that $\PiSampZ$ is computationally sound. Note that now we are in the computational setting, we cannot use the quantum de Finetti theorem as above since it only holds in the information-theoretical setting\hannote{really? maybe emphasis on constructing adversary}. Furthermore, parallel repetition for computationally sound protocols are typically difficult to analyze, and known to not always work for protocols with four or more messages~\cite{BIN97,PW12}.

% Fortunately, parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19} with interestingly different proofs. Both works showed that $m$-fold parallel repetition reduces the soundness error to $2^{-m}$ for $\BQP$ by relying on special properties of Mahadev's protocol.

 Parallel repetition of Mahadev's protocol for $\BQP$ has been analyzed before in ~\cite{arXiv:ChiaChungYam19, arXiv:AlaChiHun19}. However, the parallel repetitions of Mahadev's protocol for $\BQP$ and $\SampBQP$ is very different. For $\BQP$, the verifier simply chooses to run the Hadamard and testing rounds independently for each repetition. In contrast, our $\PiSampZ$ runs the Hadamard round in one repetition and run the testing rounds in the rest. The reason is that in $\SampBQP$, as well as generically in sampling problems, there is no known approach to combine two samples learned from two Hadamard-round executions to generate one sample with much better parameters, i.e., there is no generic error reduction method for the sampling problem. 
In contrast, the error reduction for decision problems can be done with the majority vote. 
As a result, while the soundness error decreases exponentially for $\BQP$, as we see below (and also in the above $\QPIP_1$ protocols), for $\SampBQP$, $m$-fold repetition only decreases the error to $\poly(1/m)$.

%Intuitively, Mahadev's protocol has the property that if a malicious prover $P^*$ knows the answer to pass the testing round, then it must be ``committed'' to a state $\rho$ in the sense that the verifier $V$ learns the $X/Z$ measurement outcome of $\rho$ in the Hadamard round (in a computational sense). Thus, the soundness of the underlying $\QPIP_1$ protocol ensures that $V$ learns a correct sample when it accepts. In short, when $P^*$ knows the testing round answer,  $P^*$ cannot ``cheat'' in the Hadamard round.

%\hannote{copy ch 5 here?}
To analyze the soundness of $\PiSampZ$, we use the \emph{partition lemma} developed in~\cite{arXiv:ChiaChungYam19} to analyze the prover's behavior while executing copies of $\PiMeasure$. Intuitively, the partition lemma says that for any cheating prover and for each copy $i\in[m]$, there exist two efficient ``projectors" \footnote{Actually they are not projectors, but for the simplicity of this discussion let's assume they are.} $G_{0,i}$ and $G_{1,i}$ in the prover's internal space with $G_{0,i}+G_{1,i} \approx Id$. $G_{0,i}$ and $G_{1,i}$ splits up the prover's residual internal state after sending back his first message.
$G_{0,i}$ intuitively represents the subspace where the prover does not knows the answer to the testing round on the $i$-th copy, while $G_{1,i}$ represents the subspace where the prover does. Note that the prover is using a single internal space for all copies, and every $G_{0,i}$ and every $G_{1,i}$ is acting on this single internal space. 
By using this partition lemma iteratively, we can decompose the prover's internal state $\ket{\psi}$ into sum of subnormalized states.
First we apply it to the first copy, writing $\ket{\psi}=G_{0,1}\ket{\psi}+G_{1,1}\ket{\psi} \equiv \ket{\psi_0}+\ket{\psi_1}$.
The component $\ket{\psi_0}$ would then get rejected as long as the first copy is chosen as a testing round,
which occurs with pretty high probability.
More precisely, the output corresponding to $\ket{\psi_0}$ is $1/m$-close to the ideal distribution that just rejects all the time.
On the other hand, $\ket{\psi_1}$ is now binding on the first copy;
we now similarly apply the partition lemma of the second copy to $\ket{\psi_1}$.
We write $\ket{\psi_1}=G_{0,2}\ket{\psi_1}+G_{1,2}\ket{\psi_1}\equiv \ket{\psi_{10}}+\ket{\psi_{11}}$, and apply the same argument about $\ket{\psi_{10}}$ and $\ket{\psi_{11}}$.
We then continue to decompose $\ket{\psi_{11}}=\ket{\psi_{110}}+\ket{\psi_{111}}$ and so on, until we reach the last copy and obtain $\ket{\psi_{1^m}}$.
Intuitively, all the $\ket{\psi_{1\dots10}}$ terms will be rejected with high probability, while the $\ket{\psi_{1^m}}$ term represents the ``good" component where the prover knows the answer to every testing round and therefore has high accept probability. Therefore, $\ket{\psi_{1^m}}$ also satisfies some binding property,
so the verifier should obtain a measurement result of some state on the Hadamard round copy,
and the analysis from the $\QPIP_1$ protocol $\PiSamp$ follows.

However, the intuition that $\ket{\psi_{1^m}}$ is binding to every Hadamard round is incorrect. As $G_{1,i}$ does not commute with $G_{1,j}$, $\ket{\psi_{1^m}}$ is unfortunately only binding for the $m$-th copy.
To solve this problem, we start with a pointwise argument and fix the Hadamard round on the $i$-th copy where $\ket{\psi_{1^i}}$ is binding,
and show that the corresponding output is $O(\norm{\ket{\psi_{1^{i-1}0}}})$-close to ideal.
We can later average out this error over the different choices of $i$, since not all $\norm{\ket{\psi_{1^{i-1}0}}}$ can be large at the same time. Another way to see this issue is to notice that we are partitioning a quantum state, not probability events, so there are some inconsistencies between our intuition and calculation. Indeed, the error we get in the end is $O(\sqrt{1/m})$ instead of the $O(1/m)$ we expected.

% ~\cite{arXiv:ChiaChungYam19}  introduces a \emph{partition lemma}, which roughly says that there is an efficient projection that partitions the prover's internal quantum space into a ``known-answer'' subspace $S_1$ where $P^*$ knows the answer to the testing round, and a ``not-known-answer'' subspace $S_0$ where $P^*$ does not know the answer to the testing round. For an (efficient) prover's states $\ket{\psi} \in S_1$, the verifier $V$ can produce a sound output in the Hadamard round. For (efficient) prover's states $\ket{\psi} \in S_0$, the verifier $V$ rejects in the testing round with high probability. The work of \cite{arXiv:ChiaChungYam19} used this partition lemma iteratively to each instance of the repetition to show that for a no instance $x$ for any (efficient) prover's state $\ket{\psi}$, the probability that $V$ accepts is at most roughly $2^{-m}$.

% We apply the partition lemma iteratively to the prover's state  similarly to~\cite{arXiv:ChiaChungYam19}, but with significant differences to show the soundness of $\PiSampZ$ for $\SampBQP$. In particular, extra care is required to reason about the computational indistinguishability of the verifier's output sample from the correct distribution $D_x$. To gain some intuition, let us apply the partition lemma to the prover's state $\ket{\psi}$ with respect to the first coordinate, which decompose $\ket{\psi} = \ket{\psi_0} + \ket{\psi_1}$ with $\ket{\psi_0} \in S_0$ and  $\ket{\psi_1} \in S_1$, respectively. Intuitively, the $\ket{\psi_0}$ component will be rejected whenever the first coordinate runs the testing round, but the verifier may be ``cheated'' to accepted incorrectly when the first coordinate runs the Hadamard round, which happens with probability $1/m$ and is where the error come from. For the $\ket{\psi_1}$ component, $V$ will produce a sound output if it runs the Hadamard round at the first coordinate. For the case that $V$ chooses to run the testing round at the first coordinate, we can analyze it by further applying the partition lemma to $\ket{\psi_1}$ with respect to the second coordinate, and repeat the analysis, and so on. 
% %
The intuitive analysis outlined above grosses over many technical details, and we substantiate this outline with full details in 
Section~\ref{sec:qpip0_all}.

%\hannote{put this somewhere else?}
%It is worthwhile mentioning that we came to notice some online discussion\footnote{\url{https://www.scottaaronson.com/blog/?p=3697}, e.g., comment \#25, \#26, \#42, \#48. } on the possibility of a CVQC protocol for $\SampBQP$ after we developed our own result. These comments suggested a possible reduction of $\SampBQP$ to the local Hamiltonian problem following a similar high-level idea as our solution, however, with no detail and a seemingly different technical route. More importantly, the issue that the verifier does not generate outputs in the testing round seems to be overlooked and not discussed there.

%A natural attempt to construct a CVQC protocol for $\SampBQP$ is to directly apply Mahadev's CVQC protocol for $\SampBQP$~\cite{FOCS:Mahadev18a}.  However, the requirement that the verifier needs to get a sample whenever accepting is not easily fulfilled by Mahadev's protocol. Before we explain this issue in details, let us go through a brief review of the construction and notations of Mahadev's protocol.
%\hannote{something more nature} \hannote{explain mahadev protocol} \hannote{explain how to direct apply} \hannote{qpip1}
% Let us revisit Mahadev's CVQC protocol~\cite{FOCS:Mahadev18a} first for some technical background. 
%Following~\cite{FOCS:Mahadev18a}, we formally define $\QPIP_{\tau}$
